Rapid Cycleway Prioritisation Tool

### 2020-06-09 report.Rmd
edits
More report edits

---
title: "Rapid Cycleway Prioritisation Tool"
subtitle: "Technical specification"
author: Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds
# output: bookdown::word_document2
output:
  bookdown::pdf_document2:
    toc: true
    toc_depth: 1
bibliography:
 - refs.bib
 - report-refs.bib
 - code/index.bib
---

@@

The statutory guidance on network management in response to COVID-19 outlines a range of measures that can support a safe and healthy restart, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
**This report describes the Rapid Cycleway Prioritisation Tool project, which provides aims help authorities effectively prioritise locations of new cycleways created by reallocating road space.**
<!-- Light segregation infrastructure such as flexible plastic wands (illustrated in Figure 1) and bollards can deliver a step change in provision for cycling, when implemented in combination with other measures such reduced speed limits on key corridors (as has happened on the A33 in Southampton and the A56 in Manchester) and when placed where they are most needed. -->

@@

The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) was designed to provide evidence in support of schemes eligible for the second tranche of funding.
Given the importance of acting quickly, and the unique opportunity that the crisis poses to improve physical activity levels in the long term, it is vital that local authorities have the evidence they need.
Without a nationally consistent and systematic evidence base, there is a risk that some already stretched local authorities will rely on ad-hoc methods to allocate funding.
The tool can also support authorities to make bold plans by highlighting the potential of key strategic corridors.

@@

Roads that met the spare space criteria but which were not selected as top routes were kept and can be shown in the layers named 'spare lane' and 'estimated width >10m'.
These results are hidden by default to simplify the visualisation of results and to focus attention on the most important results, including the Cohesive network.
They represent roads on which there is more than one lane in one direction, and roads with an estimated carriageway width of 10 m or more, respectively.
They can be visualised in the map by clicking on the layers button in the top lefthand corner of the map.

@@

As stated above, OpenStreetMap data on does not give comprehensive coverage in all parts of the country. 
The Cohesive network layer was developed to mitigate the disjointed nature of roads with spare space, but this layer does not include potential new routes outside the current road network.
The methodology only selects roads that have the same name for a continuous stretch of at least 500m, as candidates for top routes, meaning that some roads will be excluded from the 'top routes' result due to name changes (although most continuous road sections have consistent names).
To mitigate this limitation, roads that don't meet this criteria can still appear in the 'spare lane' and 'estimated width > 10m' layers.

>>>>

---
title: "Rapid Cycleway Prioritisation Tool"
subtitle: "Technical specification"
author: Dr Robin Lovelace & Dr Joey Talbot, Institute for Transport Studies, University of Leeds
output: bookdown::word_document2
# output:
#   bookdown::pdf_document2:
#     toc: true
#     toc_depth: 1
bibliography:
 - refs.bib
 - report-refs.bib
 - code/index.bib
---

@@

The statutory guidance on network management in response to COVID-19 outlines a range of measures that can support a safe and healthy restart, including the creation of new cycleways, reducing speed limits, widening of footways and closing roads to motor traffic via 'modal filters'.
The focus of this report is a tool that provides evidence in support of the first of these measures.
**The Rapid Cycleway Prioritisation Tool aims provide evidence on where to prioritise cycleways created by reallocating road space.**
<!-- Light segregation infrastructure such as flexible plastic wands (illustrated in Figure 1) and bollards can deliver a step change in provision for cycling, when implemented in combination with other measures such reduced speed limits on key corridors (as has happened on the A33 in Southampton and the A56 in Manchester) and when placed where they are most needed. -->

@@

The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/), supported by the Department for Transport and Sustrans, was designed to provide evidence in support of schemes eligible for the second tranche of funding.
The tool provides a national evidence base that, used alongside other sources of evidence, and help ensure that the funding is spent effectively, to maximise cycling uptake in the coming and months and years of the post COVID-19 world.

@@

Roads that met the spare space criteria but which were not selected as top routes were kept and can be shown in the layers named 'spare lane' and 'estimated width >10m'.
This includes roads with high cycling potential but which were not selected as top routes because they do not keep the same name for a stretch of at least 500 m in length.
These results are hidden by default to simplify the visualisation of results and to focus attention on the most important results, including the Cohesive network.
They represent roads on which there is more than one lane in one direction, and roads with an estimated carriageway width of 10 m or more, respectively.
They can be visualised in the map by clicking on the layers button in the top lefthand corner of the map.

@@

As stated above, OpenStreetMap data on does not give comprehensive coverage in all parts of the country. 
The Cohesive network layer was developed to mitigate the disjointed nature of roads with spare space, but this layer does not include potential new routes outside the current road network.

The methodology only selects roads that have the same name for a continuous stretch of at least 500m, as candidates for top routes, meaning that some roads will be excluded from the 'top routes' result due to name changes (although most continuous road sections have consistent names).
To mitigate this limitation, roads that don't meet this criteria can still appear in the 'spare lane' and 'estimated width > 10m' layers.
